---
title: "Coursera ML Project"
author: "Camille Roussel"
date: "June 22, 2014"
output: html_document
---
# Machine Learning Project
Based on work and data from <http://groupware.les.inf.puc-rio.br/har>

## Setup
```{r setup, message=FALSE}
library(caret)
library(randomForest)
set.seed(42)
```

## Prepare the data.
I made a decision to drop any columns that where primarily NA. I could have tried boosting these columns if the model was not accurate enough but did not see the need for this case.
```{r loaddata}
data <- read.csv('data/pml-training.csv', na.strings=c('#DIV/0!', 'NA'))

# Remove low signal data
names <- as.data.frame(sapply(data, function(x){sum(is.na(x)) / length(x)}))
names$name <- row.names(names)
names <- names[names[1] < .5,][[2]]
data <- data[,names]
data <- data[,c(-1, -3:-5)]
```

## Data slicing and sampling
The full data is too large to run on my simple laptop so we will randomly sample it to pull a reasonable subset. This subset is split into training and testing data.
```{r dataslice}
# Data slice with sampling (for reasonable run times)
data <- data[sample(1:nrow(data), 8000),]
inTrain <- createDataPartition(y=data$classe, p=.75, list=F)
data.train <- data[inTrain,]
data.test <- data[-inTrain,]
```

## Model fit
A random tree model was chosen based on the original teams work.
```{r model}
# Model
fit.rf <- train(classe~., data=data.train, method='rf', proxy=TRUE)
```

## Predictions
We use our model to run predictions against our testing data we split of earlier. We compare these results against actual.
```{r predictions}
pred <- predict(fit.rf, data.test)
data.test$correct <- pred==data.test$classe
table(pred, data.test$classe)
```

## Error Rate
The out of sample error rate from the model is presented along with our estimated out of sample error rate from our testing cross sample.
```{r OOB}
fit.rf$finalModel
paste0('Estimated OOB: ', 
       round((1-(sum(data.test$correct)/length(data.test$correct)))*100, 3)
       , '%')
```


## Results
The results of our model against the original testing data is generated for submission to the class.
```{r results}
pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("results/problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
    write.table(x,file="results/problemset.txt",quote=FALSE,row.names=FALSE,col.names=FALSE)
}

testing <- read.csv('data/pml-testing.csv', na.strings=c('#DIV/0!', 'NA'))
testing$classe <- predict(fit.rf, testing)
results <- as.character(testing$class)
pml_write_files(results)
```
